---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(dplyr)
```

```{r}

metrics = read.csv('metrics.csv')
metrics$culture_date = factor(metrics$culture_date)
cols = names(metrics)

metrics$embryo = paste(metrics$culture_date, metrics$embryo_id) 
metrics$network_id = paste(metrics$embryo, metrics$coverslip_id) 

#metrics = filter(metrics, div<=45)
```

```{r}
DT::datatable(metrics)
```
Number of recordings per condition
```{r}
aggregate(embryo ~ condition, data=metrics, FUN=length)
```
How many embryos recorded from for longer than 1 day
```{r}
a=aggregate(div ~ embryo, data=metrics, FUN=function(x) {length(unique(x))})
filter(a, div > 1)
```
Highest Power Frequency as function of DIV
```{r}
ggplot(metrics, aes(x = div, y = highest_power_freq, color = condition)) +
  geom_jitter(position=position_jitter(0.05)) +
  #geom_label(aes(label=embryo)) +
  ylab('Highest Power Frequency (Hz)') +
  xlab('DIV') + 
  theme_bw()
```
Calculate significance of differences between WT vs KO and WT vs HET
```{r}
metrics.wtvsko = filter(metrics, condition != 'mecp2het')
metrics.wtvshet = filter(metrics, condition != 'mecp2ko')

metrics.condition <- data.frame(mecp2wt_mean=numeric(),
                                mecp2wt_sd=numeric(), 
                                mecp2het_mean=numeric(), 
                                mecp2het_sd=numeric(), 
                                mecp2ko_mean=numeric(), 
                                mecp2ko_sd=numeric(), 
                                ko_pval=numeric(),
                                het_pval=numeric(),
                                stringsAsFactors=FALSE) 
metric.cols = cols[9:length(cols)]

for (col in metric.cols) {
  f = paste(col, '~ condition')
  a.mean=aggregate(formula(f), data = metrics, FUN = mean)
  rownames(a.mean) = paste(a.mean$condition, rep('_mean', 3),sep = '')
  a.std=aggregate(formula(f), data = metrics, FUN = sd)
  rownames(a.std) = paste(a.std$condition, rep('_sd', 3),sep = '')
  
  a.mean$condition = NULL
  a.std$condition = NULL
  new.metric = cbind(t(a.mean), t(a.std))
  names(new.metric) = colnames(new.metric)
  
  rko = t.test(formula(f), data=metrics.wtvsko, alternative='two.sided', var.equal=FALSE)
  rhet = t.test(formula(f), data=metrics.wtvshet, alternative='two.sided', var.equal=FALSE)
  new.metric = cbind(new.metric, data.frame(ko_pval=rko$p.value, het_val=rhet$p.value))
  
  new.metrics.condition = rbind(metrics.condition, new.metric)
  names(metrics.condition) = names(new.metric)
  rownames(new.metrics.condition) = c(rownames(metrics.condition), rownames(new.metric))
  metrics.condition = new.metrics.condition
}

```

```{r}
library(GGally)
ggscatmat(metrics, columns = 9:15, color="condition", alpha=0.8)
```
Highest power Freq by condition
```{r}

ggplot(metrics, aes(x=condition, y=highest_power_freq,color=condition)) +
  geom_boxplot() +
  ylab('Highest Power Frequency [Hz]') +
  xlab('Condition') + theme_bw() 
``` 
Mean number of events per embryo
```{r}
ggplot(metrics, aes(x = embryo, y = events_count_mean, color = condition)) +
  #geom_boxplot() +
  geom_jitter(position=position_jitter(0.05)) +
  ylab('Highest Power Frequency [Hz]') +
  xlab('Embryo') + 
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
av=aov(highest_power_freq ~ culture_date, metrics)
summary(av)

a = aggregate(highest_power_freq ~ embryo, metrics, FUN=function(x) list(mean=mean(x), sd=sd(x), count=length(x)))
DT::datatable(a)
```

Count of days with recordings from the same embryo
```{r}
embryo.count = aggregate(pcluster ~ embryo+div+condition, 
                         data=metrics,
                         FUN = function(x) c(count=length(x)))

names(embryo.count)[4] <- 'count'

embryo.div.count = aggregate(count ~ embryo + condition, data=embryo.count, FUN=length)

#ggplot(data=embryo.count, aes(x = div, y=count, color=embryo)) + 
#   geom_jitter() +
#   ylab("Count of recordings") +
#   xlab("DIV") +  theme_bw() 
```



Correlations between values
```{r}
C = cor(metrics[, c('div', cols[6:length(cols)])], method='kendall')
DT::datatable(C)
```
PCA

```{r}
#pca.cols = c('div','highest_power_freq', #'highest_oscillation_power',
#             'ipis_mean','activity_rate_mean','cluster_coeff_two')
pca.cols = c('div', cols[9:length(cols)])

#pca.dat = filter(metrics, condition != 'mecp2ko')
pca.dat = metrics

df.pca = prcomp(pca.dat[,pca.cols], center = TRUE, scale. = TRUE)
df.pca$rotation[,1:2]

library(ggbiplot)
ggbiplot(df.pca, obs.scale = 1, var.scale = 1,
              groups = pca.dat$condition, ellipse = TRUE,
              circle = FALSE, var.axes=TRUE) + 
  scale_color_discrete(name = '') + 
  theme(legend.direction = 'horizontal', legend.position = 'top')
```
```{r}
X = data.frame(df.pca$x[,seq(1,7)])
X$condition = pca.dat$condition
```


Classifier WT vs KO
```{r}
library(caret)
library(dplyr)

set.seed(825)

fitControl = trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 5)

filtered.X = filter(X, condition != 'mecp2het')
#filtered.X = X
filtered.X$condition = factor(filtered.X$condition)
#nvModel = train(condition ~ div + cluster_coeff_two + highest_power_freq + highest_oscillation_power + 
#                  ncells + activity_rate_mean + activity_rate_std + events_count_mean + events_count_std +
#                  event_ampl_mean + event_ampl_std + ipis_mean + ipis_std,
#nvModel = train(condition ~ .,
#                data = filtered.X, 
#                method = "svmLinearWeights", 
#                trControl = fitControl,
#                verbose = FALSE)
nvModel = train(condition ~ .,
                data = filtered.X, 
                method = "svmLinearWeights", 
                kmax = 3,
                trControl = fitControl,
                verbose = FALSE)
nvModel
```

Scatter plot of highest power freq and clustering coefficient as function of DIV
```{r}

ggplot(data=metrics, aes(x = div, y=highest_power_freq, color=condition)) + 
   geom_jitter(position=position_jitter(0.2)) +
   geom_text(aes(label=network_id)) +
   ylab("Highest Power Frequency") +
   xlab("DIV") +  theme_bw() 

ggplot(data=metrics, aes(x = div, y=highest_power_freq, color=condition)) + 
   geom_jitter(position=position_jitter(0.2), aes(size=highest_power_freq)) +
   ylab("Clustering strength coefficient") +
   xlab("DIV") +  theme_bw() 
```
#Variance within groups
- per embryo
```{r}
a = aggregate(highest_power_freq ~ embryo, 
          data = metrics, 
          FUN = function(x) c(mean=mean(x), var=var(x), count=length(x)))
a
```




